---
title: "A9"
author: "Elliot Gavrin"
date: "2023-03-13"
output: html_document
---
[https://github.com/19eag3/BIOL432_Assignment9](https://github.com/19eag3/BIOL432_Assignment9)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

3. 
```{r}
library(doParallel)
library(parallel)
library(foreach)
```
4. 
```{r}
N <-detectCores() 
N #number of cores
```
There are 8 cores detected.

5-6
```{r}
n_iter <- 4000
n_numbers <- 100000
mu <- 10
sigma <- 3
Time1 <-Sys.time()
for (i in 1:n_iter) {
  x <- rnorm(n = n_numbers, mean = mu, sd = sigma)
  mean_x <- mean(x)
}
Time2 <-Sys.time()
Etime <-(Time2-Time1)
print(Etime)
```

It takes 60 seconds for the for loop to complete
7. 
```{r}
S <- 1.0 # Fraction of program that is serial
speedup <- (1/((1-S) + (S/N)))
speedup
```
If the code was parallel, it would run 8 times faster than the serial code. It should take 4.275 seconds theoretically.

8. 
```{r}
Cores <- parallel::makeCluster(detectCores())
doParallel::registerDoParallel(Cores) 
n_iterations <- 4000
n_samples <- 100000
Time1 <-Sys.time()
results <- foreach(i = 1:n_iterations, .combine = c) %dopar% {
  samples <- rnorm(n_samples, mean = 10, sd = 3)
  mean(samples)
}
Time2 <-Sys.time()
parallel::stopCluster(Cores) 
Etime <-(Time2-Time1)
print(Etime)
```
The time difference of the theoretical is about 10 seconds longer than the actual run (7.4 seconds < 17.1 seconds). This is because other programs may be utilizing the cores, which causes it to not maximize all 8 cores in running the program. Because less than 8 cores are running the program, the actual time is slower than the theoretical. 



